Reproduce analyses reported in: *Phenotypic evolution is more constrained in simpler food webs*

### 2019-01-16 analyses/reproduce_analyses.Rmd
many minor edits to improve document aesthetics

```{r setup}
knitr::opts_chunk$set(messages=FALSE, warning=FALSE, error=FALSE)

# Load required libraries ----
library(car)        # for homogeneity of variance test
library(tidyverse)  # for managing data
library(cowplot)    # pretty default ggplots
library(broom)      # for tidying multiple linear models
#library(visreg)     # for quick visualizations of model effects
library(viridis)    # for color palette
library(lme4)       # for generalized linear mixed models
#library(piecewiseSEM) # for calculating model R-square
library(latex2exp)  # using Latex notation for axes labels
library(stringr)
library(MVN)        # assessing multivariate normality assumption

# Load and manage data ----

@@

  # subset data for analysis
  filter(#phenology == "early", Location == "tree",
         egg.ptoid > 0 | larval.ptoid > 0 | pupa > 0) %>%                  # eliminate unknown sources of mortality
  filter(Treatment == "Ectoparasitoid exclusion" & larval.ptoid < 1 | Treatment == "Control") %>% # excluding any larva that were parasitized by an ectoparasitoid in the exclusion experiment.
  mutate(Foodweb = ifelse(Treatment=="Control","Complex","Simple"),
         gall_survival = as.numeric(ifelse(pupa > 0, 1, 0)),
         egg.ptoid = as.numeric(ifelse(egg.ptoid > 0, 1, 0)), # there were a couple occassions where two egg parasitoids were found in a single chamber. I set these to 1 to permit an analysis with a binomial model.
         # transform and scale traits to approximate multivariate normal assumption
         sc.Diam = as.numeric(scale(chamber.diameter_mm)), 
         sc.log.Clutch = as.numeric(scale(log(clutch.size_individuals))), 
         sc.sqrt.Pref = as.numeric(scale(sqrt(oviposition.preference_individuals.per.100.shoots))))

# Custom functions to avoid repetitive code ----

@@

# Number of bootstrap replicates
n_boots_analysis <- 100
n_boots_plots <- 10

@@

```{r biased foodweb table, echo=FALSE}
# Print table
biased_foodweb_confint %>%
  select(term, estimate, conf.low, conf.high) %>%
  filter(term %in% c("FoodwebComplex:sc.Diam","FoodwebSimple:sc.Diam")) %>%
  knitr::kable(., caption = "Estimates of bias in trait-fitness relationship of chamber diameter in each food-web treatment.")
```

@@

# Tidy up estimates
tidy_foodweb_alphas <- foodweb_alphas %>%
  t() %>%
  as.data.frame() %>%
  transmute(rows = rownames(.), value=V1) %>%
  separate(col = rows, into = c("term","estimate"), sep="_") %>%
  separate(col = term, into = c("type","term_1","term_2"), sep=":", extra="drop") %>%
  mutate(term = ifelse(is.na(term_2), term_1, paste(term_1,term_2,sep=":"))) %>%
  separate(col = type, into = c("extra","type"), sep = 7) %>%
  select(type, term, estimate, value) %>%
  spread(estimate, value) %>%
  mutate(P_cutoff = ifelse(conf.high*conf.low < 0, "","*"),
         coefficient_type = ifelse(term=="(Intercept)","Mean fitness",
                                   ifelse(grepl("\\^2", .$term)==TRUE, "Quadratic",
                                          ifelse(grepl(":", .$term)==TRUE, "Correlational","Linear"))))
```
\

@@

RF_plot_foodweb_Clutch_df <- RF_Clutch$absolute_fitness %>% 
  select(-sc.Diam, -sc.sqrt.Pref) %>%
  gather(ID, absolute_fitness, -sc.log.Clutch, -Foodweb) %>%
  mutate(ID_group = ifelse(ID == "average", "average", "replicate")) %>%
  #filter(ID_group == "average") %>%
  group_by(Foodweb) %>%
  mutate(relative_fitness = absolute_fitness/mean(absolute_fitness)) %>%
  ungroup()

#RF_plot_egglarval_Clutch_df <- egglarval_RF_Clutch$absolute_fitness %>% 
#  select(-sc.Diam, -sc.sqrt.Pref) %>%
#  gather(ID, absolute_fitness, -sc.log.Clutch, -Foodweb) %>%
#  mutate(ID_group = ifelse(ID == "average", "average", "replicate")) %>%
#  mutate(Foodweb = ifelse(Foodweb=="Complex","Larval","Simple")) %>%
#  filter(Foodweb=="Larval", ID_group=="average") %>%
#  group_by(Foodweb) %>%
#  mutate(relative_fitness = absolute_fitness/mean(absolute_fitness))%>%
#  ungroup()

RF_uni_Clutch <- RF_plot_foodweb_Clutch_df %>%
#  mutate(Foodweb=factor(Foodweb, levels=c("Complex","Simple","Larval"), labels=c("Egg + Larval (Complex)","Egg (Simple)","Larval"))) %>%
#  filter(ID_group=="average") %>%
  ggplot(., aes(x=sc.log.Clutch, y=relative_fitness, group=interaction(ID,Foodweb), color=Foodweb, alpha=ID_group, size=ID_group)) +
  geom_line() +
  scale_alpha_manual(values=c(1,0.2), guide=FALSE) +
  scale_size_manual(values=c(1.5,0.25), guide=FALSE) +
  xlab("Clutch size (SDs)") +
  ylab("Relative fitness") +
  scale_color_manual(values = cbPalette[c(6,5)], name = "Food web") + 
  scale_x_continuous(breaks=c(-1,-0.5,0,0.5,1), labels = c(-1,-0.5,0,0.5,1)) +
  coord_cartesian(ylim=c(0.2,2)) +
  geom_hline(yintercept=1, linetype="dotted")

# Female pref

RF_plot_foodweb_Pref_df <- RF_Pref$absolute_fitness %>% 
  select(-sc.Diam, -sc.log.Clutch) %>%
  gather(ID, absolute_fitness, -sc.sqrt.Pref, -Foodweb) %>%
  mutate(ID_group = ifelse(ID == "average", "average", "replicate")) %>%
  #filter(ID_group == "average") %>%
  group_by(Foodweb) %>%
  mutate(relative_fitness = absolute_fitness/mean(absolute_fitness))%>%
  ungroup()

#RF_plot_egglarval_Pref_df <- egglarval_RF_Pref$absolute_fitness %>%
#  select(-sc.Diam, -sc.log.Clutch) %>%
#  gather(ID, absolute_fitness, -sc.sqrt.Pref, -Foodweb) %>%
#  mutate(ID_group = ifelse(ID == "average", "average", "replicate")) %>%
#  mutate(Foodweb = ifelse(Foodweb=="Complex","Larval","Simple")) %>%
#  filter(Foodweb=="Larval", ID_group=="average") %>%
#  group_by(Foodweb) %>%
#  mutate(relative_fitness = absolute_fitness/mean(absolute_fitness))%>%
#  ungroup()

RF_uni_Pref <- RF_plot_foodweb_Pref_df %>%
#  mutate(Foodweb=factor(Foodweb, levels=c("Complex","Simple","Larval"), labels=c("Egg + Larval (Complex)","Egg (Simple)","Larval"))) %>%
#  filter(ID_group=="average") %>%
  ggplot(., aes(x=sc.sqrt.Pref, y=relative_fitness, group=interaction(ID,Foodweb), color=Foodweb, alpha=ID_group, size=ID_group)) +
  geom_line() +
  scale_alpha_manual(values=c(1,0.2), guide=FALSE) +
  scale_size_manual(values=c(1.5,0.25), guide=FALSE) +
  xlab("Oviposition preference (SDs)") +
  ylab("Relative fitness") +
  scale_color_manual(values = cbPalette[c(6,5)], name = "Food web") + 
  scale_x_continuous(breaks=c(-1,-0.5,0,0.5,1), labels = c(-1,-0.5,0,0.5,1)) +
  coord_cartesian(ylim=c(0.2,2)) +
  theme(legend.title.align = 0.5) +
  geom_hline(yintercept=1, linetype="dotted")

# Chamber diameter
RF_plot_foodweb_Diam_df <- RF_Diam$absolute_fitness %>% 
  select(-sc.sqrt.Pref, -sc.log.Clutch) %>%
  gather(ID, absolute_fitness, -sc.Diam, -Foodweb) %>%
  mutate(ID_group = ifelse(ID == "average", "average", "replicate")) %>%
  #filter(ID_group == "average") %>%
  group_by(Foodweb) %>%
  mutate(relative_fitness = absolute_fitness/mean(absolute_fitness))%>%
  ungroup()

#RF_plot_egglarval_Diam_df <- egglarval_RF_Diam$absolute_fitness %>% 
#  select(-sc.sqrt.Pref, -sc.log.Clutch) %>%
#  gather(ID, absolute_fitness, -sc.Diam, -Foodweb) %>%
#  mutate(ID_group = ifelse(ID == "average", "average", "replicate")) %>%
#  mutate(Foodweb = ifelse(Foodweb=="Complex","Larval","Simple")) %>%
#  filter(Foodweb=="Larval", ID_group=="average") %>%
#  group_by(Foodweb) %>%
#  mutate(relative_fitness = absolute_fitness/mean(absolute_fitness))%>%
#  ungroup()

RF_uni_Diam <- RF_plot_foodweb_Diam_df %>%
#  mutate(Foodweb=factor(Foodweb, levels=c("Complex","Simple","Larval"), labels=c("Egg + Larval (Complex)","Egg (Simple)","Larval"))) %>%
#  filter(ID_group=="average") %>%
  ggplot(., aes(x=sc.Diam, y=relative_fitness, group=interaction(ID,Foodweb), color=Foodweb, alpha=ID_group, size=ID_group)) +
  geom_line() +
  scale_alpha_manual(values=c(1,0.2), guide=FALSE) +
  scale_size_manual(values=c(1.5,0.25), guide=FALSE) +
  xlab("Chamber diameter (SDs)") +
  ylab("Relative fitness") +
  scale_color_manual(values = cbPalette[c(6,5)], name = "Food web") + 
  scale_x_continuous(breaks=c(-1,-0.5,0,0.5,1), labels = c(-1,-0.5,0,0.5,1)) +
  coord_cartesian(ylim=c(0.2,2)) +
  theme(legend.title.align = 0.5) +
  geom_hline(yintercept=1, linetype="dotted")

>>>>

## Setup

Suppress messages, warnings, and errors for document aesthetics.

```{r setup}
knitr::opts_chunk$set(message=FALSE, warning=FALSE, error=FALSE)
```

Read in data, packages, and required functions.

```{r read data and functions, results='hide'}
# Load required libraries
library(car)        # for homogeneity of variance test
library(tidyverse)  # for managing data
library(cowplot)    # pretty default ggplots
library(broom)      # for tidying multiple linear models
library(viridis)    # for color palette
library(lme4)       # for generalized linear mixed models
library(latex2exp)  # using Latex notation for axes labels
library(stringr)
library(MVN)        # assessing multivariate normality assumption

# Load and manage data

@@

  # subset data for analysis
  filter(egg.ptoid > 0 | larval.ptoid > 0 | pupa > 0) %>% # eliminate unknown sources of mortality
  filter(Treatment == "Ectoparasitoid exclusion" & larval.ptoid < 1 | Treatment == "Control") %>% # excluding any larva that were parasitized by an ectoparasitoid in the exclusion experiment.
  mutate(Foodweb = ifelse(Treatment=="Control","Complex","Simple"),
         gall_survival = as.numeric(ifelse(pupa > 0, 1, 0)),
         egg.ptoid = as.numeric(ifelse(egg.ptoid > 0, 1, 0)), # there were a couple occassions where two egg parasitoids were found in a single chamber. I set these to 1 to permit an analysis with a binomial model.
         # transform and scale traits to approximate multivariate normal assumption
         sc.Diam = as.numeric(scale(chamber.diameter_mm)), 
         sc.log.Clutch = as.numeric(scale(log(clutch.size_individuals))), 
         sc.sqrt.Pref = as.numeric(scale(sqrt(oviposition.preference_individuals.per.100.shoots))))

# Custom functions and parameters to avoid repetitive code

@@

# Number of bootstrap replicates
n_boots_analysis <- 100
n_boots_plots <- 100

@@

```{r biased foodweb table, echo=FALSE}
# Print table
biased_foodweb_confint %>%
  select(term, estimate, conf.low, conf.high) %>%
  filter(term %in% c("FoodwebComplex:sc.Diam","FoodwebSimple:sc.Diam")) %>%
  mutate(estimate = round(estimate, 3), conf.low = round(conf.low, 3), conf.high = round(conf.high, 3)) %>%
  knitr::kable(., caption = "Estimates of bias in trait-fitness relationship of chamber diameter in each food-web treatment.")
```

@@

# Tidy up estimates
tidy_foodweb_alphas <- foodweb_alphas %>%
  t() %>%
  as.data.frame() %>%
  transmute(rows = rownames(.), value=V1) %>%
  separate(col = rows, into = c("term","estimate"), sep="_") %>%
  separate(col = term, into = c("type","term_1","term_2"), sep=":", extra="drop") %>%
  mutate(term = ifelse(is.na(term_2), term_1, paste(term_1,term_2,sep=":"))) %>%
  separate(col = type, into = c("extra","type"), sep = 7) %>%
  select(type, term, estimate, value) %>%
  spread(estimate, value) %>%
  mutate(P_cutoff = ifelse(conf.high*conf.low < 0, "","*"),
         coefficient_type = ifelse(term=="(Intercept)","Mean fitness",
                                   ifelse(grepl("\\^2", .$term)==TRUE, "Quadratic",
                                          ifelse(grepl(":", .$term)==TRUE, "Correlational","Linear"))))
```

@@

RF_plot_foodweb_Clutch_df <- RF_Clutch$absolute_fitness %>% 
  select(-sc.Diam, -sc.sqrt.Pref) %>%
  gather(ID, absolute_fitness, -sc.log.Clutch, -Foodweb) %>%
  mutate(ID_group = ifelse(ID == "average", "average", "replicate")) %>%
  group_by(Foodweb) %>%
  mutate(relative_fitness = absolute_fitness/mean(absolute_fitness)) %>%
  ungroup()

RF_uni_Clutch <- RF_plot_foodweb_Clutch_df %>%
  ggplot(., aes(x=sc.log.Clutch, y=relative_fitness, group=interaction(ID,Foodweb), color=Foodweb, alpha=ID_group, size=ID_group)) +
  geom_line() +
  scale_alpha_manual(values=c(1,0.2), guide=FALSE) +
  scale_size_manual(values=c(1.5,0.25), guide=FALSE) +
  xlab("Clutch size (SDs)") +
  ylab("Relative fitness") +
  scale_color_manual(values = cbPalette[c(6,5)], name = "Food web") + 
  scale_x_continuous(breaks=c(-1,-0.5,0,0.5,1), labels = c(-1,-0.5,0,0.5,1)) +
  coord_cartesian(ylim=c(0.2,2)) +
  geom_hline(yintercept=1, linetype="dotted")

# Oviposition preference

RF_plot_foodweb_Pref_df <- RF_Pref$absolute_fitness %>% 
  select(-sc.Diam, -sc.log.Clutch) %>%
  gather(ID, absolute_fitness, -sc.sqrt.Pref, -Foodweb) %>%
  mutate(ID_group = ifelse(ID == "average", "average", "replicate")) %>%
  group_by(Foodweb) %>%
  mutate(relative_fitness = absolute_fitness/mean(absolute_fitness))%>%
  ungroup()

RF_uni_Pref <- RF_plot_foodweb_Pref_df %>%
  ggplot(., aes(x=sc.sqrt.Pref, y=relative_fitness, group=interaction(ID,Foodweb), color=Foodweb, alpha=ID_group, size=ID_group)) +
  geom_line() +
  scale_alpha_manual(values=c(1,0.2), guide=FALSE) +
  scale_size_manual(values=c(1.5,0.25), guide=FALSE) +
  xlab("Oviposition preference (SDs)") +
  ylab("Relative fitness") +
  scale_color_manual(values = cbPalette[c(6,5)], name = "Food web") + 
  scale_x_continuous(breaks=c(-1,-0.5,0,0.5,1), labels = c(-1,-0.5,0,0.5,1)) +
  coord_cartesian(ylim=c(0.2,2)) +
  theme(legend.title.align = 0.5) +
  geom_hline(yintercept=1, linetype="dotted")

# Chamber diameter
RF_plot_foodweb_Diam_df <- RF_Diam$absolute_fitness %>% 
  select(-sc.sqrt.Pref, -sc.log.Clutch) %>%
  gather(ID, absolute_fitness, -sc.Diam, -Foodweb) %>%
  mutate(ID_group = ifelse(ID == "average", "average", "replicate")) %>%
  group_by(Foodweb) %>%
  mutate(relative_fitness = absolute_fitness/mean(absolute_fitness))%>%
  ungroup()

RF_uni_Diam <- RF_plot_foodweb_Diam_df %>%
  ggplot(., aes(x=sc.Diam, y=relative_fitness, group=interaction(ID,Foodweb), color=Foodweb, alpha=ID_group, size=ID_group)) +
  geom_line() +
  scale_alpha_manual(values=c(1,0.2), guide=FALSE) +
  scale_size_manual(values=c(1.5,0.25), guide=FALSE) +
  xlab("Chamber diameter (SDs)") +
  ylab("Relative fitness") +
  scale_color_manual(values = cbPalette[c(6,5)], name = "Food web") + 
  scale_x_continuous(breaks=c(-1,-0.5,0,0.5,1), labels = c(-1,-0.5,0,0.5,1)) +
  coord_cartesian(ylim=c(0.2,2)) +
  theme(legend.title.align = 0.5) +
  geom_hline(yintercept=1, linetype="dotted")
